# Age based indices


## Loading needed libraries

```{r age_indices}
library(tidyverse)
library(fjolst)    # for the time being
library(pax)
library(gam)
```

## Some older trials

Still under development (needs further testing), but here is one trial.

```{r cod_smb_age, cache=TRUE}
# Lest start from scratch
stratas <-
  husky::stratas_df %>% 
  select(strata, area = rall.area)
SPECIES <- 1
lengthclass <- c(seq(4.5, 109.5, by = 5), 119.5, 139.5)
ind <- c(31931, 31932, 32131, 36731, 37031, 37131, 37132, 37231, 41431, 41531, 42231, 42232, 47431, 52331)
st1 <-
  bind_rows(husky::STODVAR) %>%
  filter(tognumer < 20 | index %in% ind) %>%
  mutate(region = ifelse(area %in% c(1, 9, 10), "south",
                         ifelse(area %in% c(2:8), "north", NA))) %>%
  filter(!is.na(region)) %>%
  select(synis.id, ar, towlength = toglengd, region, strata = newstrata)

st2 <-
  bind_rows(husky::STODVAR) %>% 
  filter(area %in% 1:10) %>% 
  bind_rows(lesa.stodvar(leidangur="A4-2001")) %>% 
  mutate(region = ifelse(area %in% c(1, 9, 10), "south",
                         ifelse(area %in% c(2:8), "north", NA))) %>%
  filter(!is.na(region)) %>%
  select(synis.id, ar, towlength = toglengd, region, strata = newstrata)

x <- calc_age_indices(st_length = st1, st_ototliths = st2, species = 1)
```


```{r}
x$aggr %>% 
  mutate(n = round(n/1e6, 2)) %>% 
  select(-n.cv) %>% 
  filter(aldur %in% 1:11) %>% # just to fit things on the screen
  spread(aldur, n) %>% 
  as.data.frame()
```

Above it not a perfect match with the husky-approach, but who says the latter is correct :-)

## Gods code

original code on: /net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/AgeIndex/codold.r

here changed so it works irrespective of being in a specific working directory:
```{r, eval = FALSE}
TEG <- 1
home <- "/net/hafkaldi/export/u2/reikn/R/SurveyWork/"
husky::Rattach(paste(home,"ALKPrograms",sep="/"))
load("/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Stations.rdata")
lastyear <- 2018

# ------------------------------------------------------------------------------
# 1. Estimate various statistics based on data from 1993 onwards
#  This is to be used for years 1985 to 1992, a period prior to Hafvog

years <- 1993:lastyear
le <- 8:135
cn <- c("lat","lon","synis.id","area","ar")
st <- STODVAR.all[STODVAR.all$ar %in% years,cn]
tmp <- lesa.stodvar(leidangur="A4-2001")
st <- rbind(st,tmp[,c(cn)])
regs <- c("N","S") 
kvtorskur <- lesa.kvarnir(st$synis.id,TEG,c("kyn","kynthroski","slaegt","oslaegt","lifur","kynfaeri"))

#  kvtorskur <- kv.filter(kvtorskur,1) bara hægt þegar aldur er kominn.  
kvtorskur <- filter.lengdthyngd(kvtorskur,oslaegt=T)
  
#  kvtorskur <- filter.lengdthyngd(kvtorskur,oslaegt=F) # Eyðir smælki.Gert seinna.  
#  kvtorskur <- kvtorskur[!is.na(kvtorskur$lifur),]
cn <- c("synis.id","lat","lon","area","ar")
kvtorskur <- fjolst:::join(kvtorskur,st[,cn],"synis.id")
kvtorskur$reg <- ifelse(is.na(match(kvtorskur$area,c(1,9,10,13))),"N","S")
kvtorskur$kynth <- ifelse(kvtorskur$kynthroski==1 | is.na(kvtorskur$kynthroski),0,1)
  
tmp <- kvtorskur[,c("slaegt","lengd","reg","ar","oslaegt","lifur")]


tmp1 <- filter.lengdthyngd(tmp,oslaegt=F)
tmp1a <- tmp[!is.na(tmp$lifur) & tmp$lifur < 0.2*tmp$oslaegt,]
i <- tmp1a$lifur < 0.003*tmp1a$oslaegt;if(any(i)) tmp1a$lifur[i] <- tmp1a$oslaegt[i]*0.003
  
pred.data <- expand.grid(list(lengd=le,reg=regs))
pred.data$oslaegt <- pred.data$slaegt <- pred.data$lengd
x <- glm(oslaegt~log(lengd),data=tmp,family=Gamma(link=log))
x1 <- glm(slaegt~log(lengd),data=tmp1,family=Gamma(link=log))
x1a <- glm(lifur~log(lengd),data=tmp1a,family=Gamma(link=log))
  
pred.data$oslaegt <- predict(x,pred.data,type="response")
pred.data$slaegt <- predict(x1,pred.data,type="response")
pred.data$lifur <- predict(x1a,pred.data,type="response")
pred.data.first <- pred.data
  
  
tmp$wt <- rep(1,nrow(tmp)) 
pred.data$wt <- rep(0.05,nrow(pred.data))
tmp2 <- rbind(pred.data,tmp[,names(pred.data)])

tmp1$wt <- rep(1,nrow(tmp1)) 
pred.data$wt <- rep(0.05,nrow(pred.data))
tmp3 <- rbind(pred.data,tmp1[,names(pred.data)])
  
tmp1a$wt <- rep(1,nrow(tmp1a)) 
pred.data$wt <- rep(0.05,nrow(pred.data))
tmp4 <- rbind(pred.data,tmp1a[,names(pred.data)])
  
x <- gam(oslaegt~s(log(lengd),df=8)*factor(reg),data=tmp2,family=Gamma(link=log))
i <- tmp2$wt < 0.1
pred.data$oslaegt <- x$fitted.values[i]
  
x1 <- gam(slaegt~s(log(lengd),df=8)*factor(reg),data=tmp3,family=Gamma(link=log))
i <- tmp3$wt < 0.1
pred.data$slaegt <- x1$fitted.values[i]
  
x1a <- gam(lifur~s(log(lengd),df=8)*factor(reg),data=tmp4,family=Gamma(link=log))
i <- tmp4$wt < 0.1
pred.data$lifur <- x1a$fitted.values[i]





x3 <- apply.shrink(tmp1$slaegt, tmp1$lengd, mean,names=c("lengd","slaegt.maelt"))
x3 <- fjolst:::join(x3,apply.shrink(tmp$oslaegt, tmp$lengd, mean,names=c("lengd","oslaegt.maelt")),"lengd")
x3 <- fjolst:::join(x3,apply.shrink(tmp1a$lifur, tmp1a$lengd, mean,names=c("lengd","lifur.maelt")),"lengd")

pred.data <- fjolst:::join(pred.data,x3,"lengd")
pred.data$oslaegt1 <- pred.data.first$oslaegt # úr a*l^b
pred.data$slaegt1 <- pred.data.first$slaegt # úr a*l^b
pred.data$lifur1 <- pred.data.first$lifur # úr a*l^b

torskur.marsrall.wt.all <- pred.data
  
#  par(ask=T)
#  plot(pred.data$lengd, pred.data$oslaegt.maelt/pred.data$oslaegt)
#  plot(log(pred.data$lengd),log(pred.data$oslaegt))  
#  plot(pred.data$lengd, pred.data$slaegt.maelt/pred.data$slaegt)  
#  plot(log(pred.data$lengd),log(pred.data$slaegt))  
#  plot(pred.data$lengd, pred.data$lifur.maelt/pred.data$lifur)  
#  plot(log(pred.data$lengd),log(pred.data$lifur))  

pred.data <- list(ar=years,reg=c("N","S"),lengd=le)
pred.data <- expand.grid(pred.data)
pred.data$reg <- as.character(pred.data$reg)
pred.data <- merge(pred.data,torskur.marsrall.wt.all[,c("lengd","reg","oslaegt","slaegt","lifur")])
pred.data$wt <- rep(0.1,nrow(pred.data)) 
cn <- names(pred.data)

tmp <- filter.lengdthyngd(kvtorskur,oslaegt=F)
tmp$wt <- rep(1,nrow(tmp)) #*kvtorskur$rat
tmp <- rbind(tmp[,cn],pred.data[,cn])

for( ar in years) {
    print(ar)
    tmp1 <- tmp[tmp$ar==ar ,]
    tmp3 <- tmp1
    x <- gam(slaegt~s(log(lengd),df=7)*factor(reg),data=tmp1,family=Gamma(link=log),weight=tmp1$wt,maxit=100)
    x1 <- predict(x,se.fit=T,type="response")
    j <- tmp1$wt < 1
    tmp1 <- tmp1[j,c("ar","lengd","reg")]
    tmp1$slaegt <- x$fitted.values[j]
    tmp1$slaegt.se <- x1$se.fit[j]
    if(ar==1993)
        tmp2 <- tmp1
    else
        tmp2 <- rbind(tmp2,tmp1)
}
tmpsl <- tmp2  

# Óslægt
cat("oslaegt")
  
tmp <- filter.lengdthyngd(kvtorskur,oslaegt=T)
tmp$wt <- rep(1,nrow(tmp)) #*kvtorskur$rat
tmp <- rbind(tmp[,cn],pred.data[,cn])

for( ar in years) {
    print(ar)
    tmp1 <- tmp[tmp$ar==ar ,]
    tmp3 <- tmp1
    x <- gam(oslaegt~s(log(lengd),df=7)*factor(reg),data=tmp1,family=Gamma(link=log),weight=tmp1$wt,maxit=100)
    x1 <- predict(x,se.fit=T,type="response")
    j <- tmp1$wt < 1
    tmp1 <- tmp1[j,c("ar","lengd","reg")]
    tmp1$oslaegt <- x$fitted.values[j]
    tmp1$oslaegt.se <- x1$se.fit[j]
    if(ar==1993)
        tmp2 <- tmp1
    else
        tmp2 <- rbind(tmp2,tmp1)
}
tmposl <- tmp2


cat("lifur") 
  
tmp <- kvtorskur[!is.na(kvtorskur$lifur) & kvtorskur$lifur < 0.2*kvtorskur$oslaegt,]
i <- tmp$lifur < 0.003*tmp$oslaegt;if(any(i)) tmp$lifur[i] <- tmp$oslaegt[i]*0.003
tmp <- tmp[tmp$lengd >= 25,] # Marklaust að vera með lifur fyrir minni
tmp$wt <- rep(1,nrow(tmp)) #*kvtorskur$rat
tmp <- rbind(tmp[,cn],pred.data[,cn])
for( ar in years) {
    print(ar)
    tmp1 <- tmp[tmp$ar==ar ,]
    tmp3 <- tmp1
    x <- gam(lifur~s(log(lengd),df=7)*factor(reg),data=tmp1,family=Gamma(link=log),weight=tmp1$wt,maxit=100)
    x1 <- predict(x,se.fit=T,type="response")
    j <- tmp1$wt < 1
    tmp1 <- tmp1[j,c("ar","lengd","reg")]
    tmp1$lifur <- x$fitted.values[j]
    tmp1$lifur.se <- x1$se.fit[j]
    if(ar==1993)
        tmp2 <- tmp1
    else
        tmp2 <- rbind(tmp2,tmp1)
}
tmpli <- tmp2

pred.data <- merge(pred.data[,c("ar","reg","lengd")],tmpsl)
pred.data <- merge(pred.data,tmposl)
pred.data <- merge(pred.data,tmpli)
  
codmarchsurwt <- pred.data

codmarchsurwt.8592 <- apply.shrink(codmarchsurwt$oslaegt,list(codmarchsurwt$lengd,codmarchsurwt$reg),mean,names=c("lengd","reg","oslaegt"))
#Lwt relationship finished. 

# ------------------------------------------------------------------------------
# 2. Age index
years <- 1985:lastyear
 

husky::Rattach(paste(home,"OldStratas",sep=""))
ind <- c(31931,31932,32131,36731,37031,37131,37132,37231,41431,41531,42231,42232,47431,52331)
notfixedInd <- c(27401,37212,37302,41214,41412,46211,46212,46214,46216,46311,46312,46313,51301,52413,56214,57412,62311,71912,72314)
fj  <- list()

stcol <- c("synis.id","lat","lon","oldstrata","newstrata","area","toglengd") # to use in MakeLdistbystation
regs <- list(S=c(1,9:10),N=2:8)
regname <- names(regs)
aldur <- 1:14
lengd <- c(seq(4.5,109.5,by=5),119.5,139.5)

# Calculate the age indices
for( i in 1:length(years) ) {
  print(i) 
  fj[[i]] <- data.frame()
  if(i == 17)  stauka <- lesa.stodvar(leidangur="A4-2001")
  # For each region (South and North)
  for( j in 1:2) {
    st <- STODVAR.all[!is.na(match(STODVAR.all$area,regs[[j]])) & (STODVAR.all$tognumer < 20 | !is.na(match(STODVAR.all$index,ind))) & STODVAR.all$ar==years[i] ,]
    st.all <- STODVAR.all[!is.na(match(STODVAR.all$area,regs[[j]])) & STODVAR.all$ar==years[i],]
    st <- st[is.na(match(st$index,notfixedInd)),]
    
    # If the year is 2001, add cruise A4-2001
    if(i == 17){
      cn <- c("lat","lon","synis.id")
      sttmp <- stauka[stauka$area %in% regs[[j]],cn]
      st.all <- rbind(st.all[,cn],sttmp[,cn])
    }
    
    le <- lesa.lengdir(st$synis.id,TEG)
    kv <- lesa.kvarnir(st.all$synis.id,TEG,col.names=c("kyn","kynthroski"),oracle=F)
    
    nu <- lesa.numer(st$synis.id,TEG)
    
    alk <- MakeAlk(kv, TEG, lengd=lengd, aldur = aldur, FilterAldurLengd = F, kynth = T) # Tók filteraldurlengd af
    ar <- years[i]
    if(ar < 1993){ 
      tmp <- codmarchsurwt.8592[codmarchsurwt.8592$reg==regname[j],]}
    else {
      tmp <- codmarchsurwt[codmarchsurwt$reg==regname[j] & codmarchsurwt$ar==ar,]}
    ldist <- MakeLdistbyStation(le,nu,1,lengd=lengd,Stodvar=st,talid=T,lengd.thyngd.data=tmp,stodvar.col=stcol)
    fj[[i]]  <- rbind.alk(fj[[i]],Calc.fj.per.station(alk,ldist))
  }
}

torskur.fj <- fj
names(torskur.fj) <- as.character(years)



# Aggregate statistics

biomass <- kynthbiomass <-  meanlefj <- sdevfj <- kynthfj <- fj <- list()
biovisit <- fjvisit <- kynthvisit <- meanlevisit <-  sdevvisit <- kynthbiovisit  <- list()

for(i in 1:length(years)) {
  print(i)
  st <- attributes(torskur.fj[[i]])$Stodvar
  fj <- kynthfj <- biomass <- meanlefj <- sdevfj <- kynthbiomass <- list()
  biom <- torskur.fj[[i]]$WtPerAldur*torskur.fj[[i]]$FjPerAldur
  biom[is.na(biom)] <- 0
  kynthbiom <- torskur.fj[[i]]$KynthWtPerAldur*torskur.fj[[i]]$KynthFjPerAldur
  kynthbiom[is.na(kynthbiom)] <- 0
  # For each age group
  for(j in 1:14) {
    cat(paste(j," "))
    fj[[j]] <- Calc.index(st,z=torskur.fj[[i]]$FjPerAldur[,j],cn="oldstrata")$aggr.output
    kynthfj[[j]] <- Calc.index(st,z=torskur.fj[[i]]$KynthFjPerAldur[,j],cn="oldstrata")$aggr.output
    biomass[[j]] <- Calc.index(st,z=biom[,j],cn="oldstrata")$aggr.output
    meanlefj[[j]] <- Calc.index(st,z=torskur.fj[[i]]$LengdSinnumFjPerAldur[,j],cn="oldstrata")$aggr.output
    sdevfj[[j]] <- Calc.index(st,z=torskur.fj[[i]]$Lengd2SinnumFjPerAldur[,j],cn="oldstrata")$aggr.output
    kynthbiomass[[j]] <- Calc.index(st,z=kynthbiom[,j],cn="oldstrata")$aggr.output
  }
  biovisit[[i]] <- biomass
  fjvisit[[i]] <- fj
  kynthvisit[[i]] <- kynthfj
  kynthbiovisit[[i]] <- kynthbiomass
  sdevvisit[[i]] <- sdevfj
  meanlevisit[[i]] <- meanlefj
}
names(fjvisit) <- names(biovisit) <- names(meanlevisit) <- names(sdevvisit) <-names(kynthvisit) <- names(kynthbiovisit) <-  as.character(years)


codindexlistn <- combine.alk.visit(fjvisit, kynthvisit, biovisit,meanlevisit,sdevvisit,kynthbiovisit ,row = 13, aldur = 1:14,ar=years)
codindexlists <- combine.alk.visit(fjvisit, kynthvisit, biovisit,meanlevisit,sdevvisit,kynthbiovisit,row = c(14), aldur = 1:14,ar=years)
codindexlistanF  <- combine.alk.visit(fjvisit, kynthvisit, biovisit, meanlevisit,sdevvisit,kynthbiovisit,row = c(22), aldur = 1:14,ar=years)


codindexlist <- combine.alk.visit(fjvisit, kynthvisit, biovisit,meanlevisit,sdevvisit,kynthbiovisit ,row = 15, aldur = 1:14,ar=years)
#torskur.visit.faer  <- combine.alk.visit(fjvisit, kynthvisit, biovisit,meanlevisit,sdevvisit,kynthbiovisit ,row = 7, aldur = 1:14,ar=years)

x <- codindexlist$fj - codindexlistanF$fj
x <- x[as.character(1994:1995),]
tmp <- codindexlistanF$fj
tmp <- tmp[as.character(1994:1995),]
x1 <- x/(tmp+1e-9)
faerhrrat9495 <- apply(x1, 2, mean)


tmp <- codindexlist$fj
for(i in 1996:2003)
        tmp[as.character(i),] <- tmp[as.character(i),] * (1 + faerhrrat9495)
codindexlist$fj <- tmp



# Setja saman fyrir ORACLE

age <- 1:14
year <- years
n <- length(age)
age <- matrix(age,n,length(year)) 
year <- matrix(year,n,length(year),byrow=T)
yearage <- data.frame(year=c(year),age=c(age))

tmp <- tmp1 <- tmp2 <- tmp3 <-  yearage
tmp$reg <- "Tot"
tmp$fj <- c(t(codindexlist$fj))
tmp$cv <- c(t(codindexlist$cv))
tmp$wt <- c(t(codindexlist$wt))
tmp$kynthhlutfall <- c(t(codindexlist$kynthhlutfall))
tmp$meanle <- c(t(codindexlist$meanle))
tmp$sdev <- c(t(codindexlist$sdev))
tmp$kynthwt <- c(t(codindexlist$kynthwt))

tmp1$reg <- "N"
tmp1$fj <- c(t(codindexlistn$fj))
tmp1$cv <- c(t(codindexlistn$cv))
tmp1$wt <- c(t(codindexlistn$wt))
tmp1$kynthhlutfall <- c(t(codindexlistn$kynthhlutfall))
tmp1$meanle <- c(t(codindexlistn$meanle))
tmp1$sdev <- c(t(codindexlistn$sdev))
tmp1$kynthwt <- c(t(codindexlistn$kynthwt))

tmp2$reg <- "S"
tmp2$fj <- c(t(codindexlists$fj))
tmp2$cv <- c(t(codindexlists$cv))
tmp2$wt <- c(t(codindexlists$wt))
tmp2$kynthhlutfall <- c(t(codindexlists$kynthhlutfall))
tmp2$meanle <- c(t(codindexlists$meanle))
tmp2$sdev <- c(t(codindexlists$sdev))
tmp2$kynthwt <- c(t(codindexlists$kynthwt))

tmp3$reg <- "TotanFaer"
tmp3$fj <- c(t(codindexlistanF$fj))
tmp3$cv <- c(t(codindexlistanF$cv))
tmp3$wt <- c(t(codindexlistanF$wt))
tmp3$kynthhlutfall <- c(t(codindexlistanF$kynthhlutfall))
tmp3$meanle <- c(t(codindexlistanF$meanle))
tmp3$sdev <- c(t(codindexlistanF$sdev))
tmp3$kynthwt <- c(t(codindexlistanF$kynthwt))





codindextable <- rbind(tmp,tmp1,tmp2,tmp3)


codindexlist$fj <- t(round(codindexlist$fj,2))
codindexlist$cv <- t(round(codindexlist$cv,3))
codindexlist$kynthhlutfall<- t(round(codindexlist$kynthhlutfall*100,1))
codindexlist$wt <- t(round(codindexlist$wt*1000))
codindexlist$kynthwt <- t(round(codindexlist$kynthwt*1000))
codindexlist$meanle <- t(round(codindexlist$meanle,1))
codindexlist$sdev <- t(round(codindexlist$sdev,1))

codindextable$fj <- round(codindextable$fj,2)
codindextable$cv <- round(codindextable$cv,3)
codindextable$kynthhlutfall<- round(codindextable$kynthhlutfall*100,1)
codindextable$wt <- round(codindextable$wt*1000)
codindextable$kynthwt <- round(codindextable$kynthwt*1000)
codindextable$meanle <- round(codindextable$meanle,1)
codindextable$sdev <- round(codindextable$sdev,1)


#save(list=c("codmarchsurwt","codmarchsurwt.8592","codindextable","codindexlist"),file="codindexold.rdata")


```

## Reconstructing the stuff

### The stations used

```{r, eval = FALSE}
load("/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Stations.rdata")
st <- STODVAR.all[STODVAR.all$ar %in% years,]
tmp <- lesa.stodvar(leidangur="A4-2001")
ind <- c(31931,31932,32131,36731,37031,37131,37132,37231,41431,41531,42231,42232,47431,52331)
notfixedInd <- c(27401,37212,37302,41214,41412,46211,46212,46214,46216,46311,46312,46313,51301,52413,56214,57412,62311,71912,72314)
```

