# Length based indices via mar

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```


__The flow__:
```
Stations %>%
trim towlength %>%                 # narrow extremes
gather length data %>%
scale by counted %>% 
standardize by towlength %>%       # on the number only
calculate_biomass %>%              # up to now on each length class
summarise_by_station %>%           # now one record per station
                                   # Note statistic up to now is equivalent
                                   #   to afli per tog
                                   # HERE: filter station to use
                                   #       i.e. fixed etc
summarise_by_strata %>%            # one record per strata
summarise                          # one record per year
```

__TO DO__:

* Indices by length class
* Indices when no length measurements, only counts. Possibly use full_join and then set length class to zero where length class not defined.
* User defined grouping?

__Loading needed libraries__:

```{r length_indices_mar}
library(tidyverse)
library(mar)
```

__Some functions__:
```{r}
trim_towlength <- function(d, std.towlength = 4, min.towlength = 2, max.towlength = 8) {
  
  d %>% 
    # seems like one can not refer to same variable twice within one single mutate
    mutate(toglengd = if_else(toglengd > max.towlength, max.towlength, toglengd)) %>% 
    mutate(toglengd = if_else(toglengd < min.towlength, min.towlength, toglengd))
}


gather_length_data <- function(Stations, Species, Length = 0) {
  
  # Here could create an switch, if Stations is connection do below
  #   else read from fjolst
  Lengths <- 
    lesa_lengdir(Stations$src) %>% 
    filter(tegund %in% Species,
           lengd >= Length) %>% 
    group_by(synis_id, tegund, lengd) %>% 
    summarise(fjoldi = sum(fjoldi, na.rm = TRUE)) %>% 
    ungroup()
  
  Counts <-
    lesa_numer(Stations$src) %>% 
    filter(tegund %in% Species) %>% 
    mutate(r =  ifelse(fj_maelt != 0, 1 + fj_talid/fj_maelt, 1)) %>% 
    select(synis_id, tegund, fj_maelt, fj_talid, r)
  
  Stations %>% 
    left_join(Lengths, by = "synis_id") %>% 
    left_join(Counts, by = c("synis_id", "tegund"))
  
}


scale_by_counted <- function(d) {
  d %>% 
    mutate(fjoldi = r * fjoldi / 1e3) %>%   # units of thousan
    select(-r)                              # no use anymore
}

standardise_towlength <- function(d, std.towlength = 4) {
  
  d %>% 
    mutate(N = fjoldi * std.towlength/toglengd) %>%   # standardize to per 4 miles
    select(-toglengd)

}

calculate_biomass <- function(d) {
  
  lwcoeff <- tbl_mar(con, "ops$einarhj.lwcoeff")
  
  d %>% 
    left_join(lwcoeff, by = "tegund") %>% 
    # use Newton's law if lwcoefficient for species not specified
    mutate(a = ifelse(is.na(a), 0.01, a),
           b = ifelse(is.na(b), 3.00, b)) %>% 
    # Below statistic is by length class
    # If include lwcoeff then do a priori a left_join
    mutate(B  = ifelse(is.na(N), 0, N) * a * lengd^b/1e3) %>% 
    select(-a, -b)
  
}

calc_by_strata <- function(d) {
  
  # Stuff should not really be inside this function
  Stratas <- 
  tbl_mar(con, "ops$einarhj.stratas") %>% 
  select(strata, area = rall.area)
    
  d %>% 
    group_by(ar, strata) %>% 
  dplyr::summarise(sN  = n(),   # Number of stations within strata
                   n_m  = mean(N, na.rm = TRUE),
                   n_d  = ifelse(n() == 1, mean(N, na.rm = TRUE) * std.cv, sd(N)),
                   b_m  = mean(B, na.rm = TRUE),
                   b_d  = ifelse(n() == 1, mean(B, na.rm = TRUE) * std.cv, sd(B))) %>% 
  #b_d  = ifelse(N == 1, b_m  * std.cv, stats::sd(b)),
  ungroup() %>%
  left_join(Stratas, by = "strata") %>%
  # within oracle, have to mutate in stages
  mutate(area  = area/1.852^2 / std.area) %>% 
  mutate(n     = n_m  * area,
         b     = b_m  * area)
  
}

aggregate_by_year <- function(d) {
  
  # below equvialent to aggr-calculation
  d %>% 
    group_by(ar) %>%
    # A la Höski:
    summarise(n = sum(n, na.rm = TRUE),
              n.cv = pax:::calc_cv(n_m, n_d, area, sN),
              b = sum(b, na.rm = TRUE),
              b.cv = pax:::calc_cv(b_m, b_d, area, sN)) %>% 
    ungroup()
  
}
```

__Run the flow__:
```{r}
LENGTH <- 5 # Index for fish length equal to or larger than
con <- connect_mar()
```

**Let's clean up the lates mess from the vatican**:
```{r, eval = FALSE}
## from /net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Indices.r
## Fixed stations not originally set out
#fixedind <- c(31931, 31932, 32131, 36731, 37031, 37131, 37132,
#              37231, 41431, 41531, 42231, 42232, 47431, 52331)
## wat!
#notfixedInd <- c(27401, 37212, 37302, 41214, 41412, 46211, 46212,
#                 46214, 46216, 46311, 46312, 46313, 51301, 52413,

# NOT RUN
attach("/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Stations.rdata")
smb <-
  STODVAR.all %>% 
  filter(tognumer %in% 1:39 | is.na(tognumer)) %>% 
  select(synis_id = synis.id, ar, toglengd, strata = oldstrata, lon, lat)
dbWriteTable(con, name = "smbstations", value = smb, overwrite = TRUE)
detach("file:/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Stations.rdata")
```


```{r}
Stations <-  
  tbl(con, "smbstations")
std.cv <- 1
std.area <- 4 * 17/1852
Stations %>%
  trim_towlength() %>% 
  #left_join(Stratas) %>% 
  gather_length_data(Species = 1, Length = LENGTH) %>%
  scale_by_counted() %>% 
  standardise_towlength() %>% 
  calculate_biomass() %>%
  # since not all length classes included can only do the total
  # for the time being to a "verb"
  group_by(synis_id, ar, strata, tegund) %>% 
  summarise(N = sum(N, na.rm = TRUE),
            B = sum(B, na.rm = TRUE)) %>% 
  # Zero stations
  mutate(N = ifelse(is.na(N), 0, N),
         B = ifelse(is.na(B), 0, B)) %>% 
  calc_by_strata() %>% 
  # Have to collect here because of calc_cv function in the aggregate step
  collect(n = Inf) %>% 
  aggregate_by_year() ->
  d
```

__A double test__:
```{r}
p <- 
  d %>% 
  select(year = ar, cb = b, cb.cv = b.cv) %>% 
  ggplot(aes(year, cb)) +
  geom_pointrange(aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "red", lwd = 1) +
  geom_line(colour = "red", lwd = 1) +
  scale_colour_brewer(palette = "Set1")


attach("/net/hafkaldi/export/u2/reikn/Splus5/SMB/Allindices.RData")
mri.2016 <- 
  All.indices %>% 
  filter(species == 1,
         svaedi == "Heild",
         lengd == LENGTH,
         type == "all") %>% 
  select(year = ar, cb = bio.staerri, cb.cv = cv.bio.staerri, type) %>% 
  mutate(source = "mri.2016")
detach("file:/net/hafkaldi/export/u2/reikn/Splus5/SMB/Allindices.RData")





attach("/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Allaggroldsmbindex.rdata")
mri <- 
  Allaggroldsmbindex %>% 
  filter(species == 1,
         svaedi == "Heild",
         lengd == LENGTH,
         diurnal == 0,
         fixed == 0) %>% 
  select(year = ar, cb = bio.staerri, cb.cv = cv.bio.staerri) %>% 
  mutate(source = "mri") %>% 
  as_tibble()
detach("file:/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Allaggroldsmbindex.rdata")
p +
  geom_pointrange(data = mri, 
                  aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "yellow") +
  geom_line(data = mri, colour = "yellow") + 
  geom_pointrange(data = mri.2016,
                  aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "blue") +
  labs(x = NULL, y = NULL,
       title = "Biomass indices",
       subtitle = "Comparison of the mar (red) mar 2016 (blue) and the Bible (grey)")
```

__Digression - calculate only those stations "now" taken__:

```{r}
con <- connect_mar()
```

```{r, eval = FALSE}
# not run
st <- 
  tbl_mar(con, "ops$einarhj.smb19852016") %>% 
  left_join(lesa_stodvar(con) %>% select(synis_id, reitur, tognumer,
                                        lon = kastad_v_lengd,
                                        lat = kastad_n_breidd)) %>% 
  collect() %>% 
  mutate(id = paste(reitur, tognumer))

st17 <- 
  lesa_stodvar(con) %>% 
  filter(ar == 2017, synaflokkur == 30, veidarfaeri == 73) %>% 
  select(synis_id, ar, reitur, toglengd,
         lon = kastad_v_lengd,
         lat = kastad_n_breidd,
         tognumer) %>% 
  collect() %>% 
  mutate(id = paste(reitur, tognumer))

# Note need to put strata id into the 2017 data
st17 <-
  st17 %>% 
  left_join(st %>% 
  filter(ar == 2016) %>% 
  select(id, strata))

STTAKEN4 <-
  st %>% 
  filter(st$id %in% st17$id) %>% 
  bind_rows(st17) %>% 
  filter(!is.na(strata))

copy_to(con, STTAKEN4, temporary = FALSE)
```

```{r}
calc_by_strata <- function(d) {
  
  # Stuff should not really be inside this function
  Stratas <- 
  tbl_mar(con, "ops$einarhj.stratas") %>% 
  select(strata, area = rall.area)
    
  d %>% 
    group_by(ar, strata, tegund) %>% 
  dplyr::summarise(sN  = n(),   # Number of stations within strata
                   n_m  = mean(N),
                   n_d  = ifelse(n() == 1, mean(N) * std.cv, sd(N)),
                   b_m  = mean(B),
                   b_d  = ifelse(n() == 1, mean(B) * std.cv, sd(B))) %>% 
  #b_d  = ifelse(N == 1, b_m  * std.cv, stats::sd(b)),
  ungroup() %>%
  left_join(Stratas, by = "strata") %>%
  # within oracle, have to mutate in stages
  mutate(area  = area/1.852^2 / std.area) %>% 
  mutate(n     = n_m  * area,
         b     = b_m  * area)
  
}

calc_by_strata <- function(d) {
  
  # Stuff should not really be inside this function
  Stratas <- 
  tbl_mar(con, "ops$einarhj.stratas") %>% 
  select(strata, area = rall.area)
    
  d %>% 
    group_by(ar, strata, species) %>% 
  dplyr::summarise(sN  = n(),   # Number of stations within strata
                   n_m  = mean(N),
                   n_d  = ifelse(n() == 1, mean(N) * std.cv, sd(N)),
                   b_m  = mean(B),
                   b_d  = ifelse(n() == 1, mean(B) * std.cv, sd(B))) %>% 
  #b_d  = ifelse(N == 1, b_m  * std.cv, stats::sd(b)),
  ungroup() %>%
  left_join(Stratas, by = "strata") %>%
  # within oracle, have to mutate in stages
  mutate(area  = area/1.852^2 / std.area) %>% 
  mutate(n     = n_m  * area,
         b     = b_m  * area)
  
}

aggregate_by_year <- function(d) {
  
  # below equvialent to aggr-calculation
  d %>% 
    group_by(ar, species) %>%
    # A la Höski:
    summarise(n = sum(n),
              n.cv = pax:::calc_cv(n_m, n_d, area, sN),
              b = sum(b),
              b.cv = pax:::calc_cv(b_m, b_d, area, sN)) %>% 
    ungroup()
  
}

Length <- 5 # Index for fish length equal to or larger than
con <- connect_mar()
Stations <-  tbl_mar(con, "ops$einarhj.STTAKEN4")

teg <- lesa_tegundir(con) %>% 
  collect(n = Inf) %>% 
  arrange(tegund)

teg2 <-
  teg %>%
  filter(tegund %in% c(998, 53, 72, 4))

std.cv <- 1
std.area <- 4 * 17/1852
Stations %>%
  trim_towlength() %>% 
  #left_join(Stratas) %>% 
  gather_length_data(Species = teg2$tegund, Length = Length) %>%
  scale_by_counted() %>% 
  standardise_towlength() %>% 
  calculate_biomass() %>%
  # since not all length classes included can only do the total
  # for the time being to a "verb"
  group_by(synis_id, ar, strata, tegund) %>% 
  summarise(N = sum(N),
            B = sum(B)) %>% 
  # Zero stations
  mutate(N = ifelse(is.na(N), 0, N),
         B = ifelse(is.na(B), 0, B)) %>% 
  rename(species = tegund) %>% 
  calc_by_strata() %>% 
  # Have to collect here because of calc_cv function in the aggregate step
  collect(n = Inf) %>% 
  aggregate_by_year() ->
  d
d %>% 
  filter(!is.na(species)) %>% 
  mutate(n = n/1e3) %>% 
  select(year = ar, y = n, cv = n.cv, species) %>% 
  ggplot(aes(year, y)) +
  geom_pointrange(aes(ymin = y * (1 - cv),
                      ymax = y * (1 + cv)),
                  colour = "red", lwd = 1) +
  geom_line(colour = "red", lwd = 1) +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ species, scale = "free_y") +
  labs(x = NULL, y = NULL)
```

```{r}
con <- connect_mar()
st <- 
  lesa_stodvar(con) %>% 
  filter(ar == 2017, synaflokkur == 30, veidarfaeri == 73) %>% 
  select(skip,
         synis_id, ar, reitur, toglengd,
         lon = kastad_v_lengd,
         lat = kastad_n_breidd,
         tognumer)
nu <-
  st %>% 
  select(synis_id) %>% 
  left_join(lesa_numer(con)) %>% 
  collect(n = Inf)
le <- 
  st %>% 
  select(synis_id) %>% 
  left_join(lesa_lengdir(con)) %>% 
  collect(n = Inf)
st <-
  st %>% 
  collect(n = Inf)
nu <-
  nu %>% 
  mutate(r = (fj_talid + fj_maelt)/ifelse(fj_maelt == 0, 1, fj_maelt))

bio <-
  le %>% 
  left_join(nu %>% select(synis_id, tegund, r)) %>% 
  mutate(fjoldi = r  * fjoldi,
         bio = fjoldi * 0.01 * lengd^3) %>% 
  group_by(synis_id) %>% 
  summarise(bio = sum(bio, na.rm = TRUE)/1e3)
n.tegundir <-
  nu %>% 
  group_by(synis_id) %>% 
  summarise(n.tegundir = n_distinct(tegund))
nu2 <-
  nu %>% 
  select(synis_id, fj_maelt, fj_talid, fj_maelt, fj_magasyna) %>% 
  gather(key = key, value = n, -synis_id) %>% 
  group_by(synis_id, key) %>% 
  summarise(n = sum(n)) %>% 
  spread(key, n)
st %>% 
  left_join(bio) %>% 
  left_join(nu2) %>%
  left_join(n.tegundir) %>% 
  group_by(skip) %>% 
  summarise(n.stodvar = n(),
            n.tegundir = mean(n.tegundir),
            bio.min = min(bio),
            bio.max = max(bio),
            bio.mean = mean(bio),
            bio.median = median(bio),
            fj_maelt = median(fj_maelt),
            fj_talid = median(fj_talid),
            fj_magasyna = median(fj_magasyna)) %>% 
  gather(key = key, value = value, -skip) %>% 
  spread(key = skip, value = value)

```

