# Length based indices

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Loading needed libraries
___

```{r length_indices}
library(tidyverse)
library(fjolst)    # for the time being
library(pax)
```

## Length based indices
___

### Specify the stations and strata

OK, here you have to have some background knowledge. But we select "all" SMB stations:

```{r}
Station <- 
  husky::STODVAR %>% 
  bind_rows() %>% 
  filter(tognumer %in% 1:39) %>% 
  select(id = synis.id, year = ar, towlength = toglengd, strata = newstrata) %>% 
  mutate(towlength = trim_towlength(towlength),
         mult = 1) %>% 
  arrange(id)
```

And get the appropriate strata as well (the old strata, which is not the same as the very old strata):

```{r}
Stratas <-
  husky::stratas_df %>% 
  select(strata, area = rall.area)
```

And then we leave the rest to the `smx:calc_length_indices`-function:

#### Cod spring survey index

```{r cod_smb_length, cache = TRUE}
res <- pax::calc_length_indices(Station, Stratas, SPECIES = 1)
```

Extracting the tidy information and preparing the plot:
```{r}
tidy_fixed <-
  res$aggr %>% 
  filter(length == 5) %>% 
  mutate(source = "tidy")
p <- 
  tidy_fixed %>% 
  select(year, cb, cb.cv) %>% 
  ggplot(aes(year, cb)) +
  geom_pointrange(aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "red", lwd = 1) +
  geom_line(colour = "red", lwd = 1) +
  scale_colour_brewer(palette = "Set1")
```

Lets add the official calculations for comparison:
```{r, plot1}
attach("/net/hafkaldi/export/u2/reikn/Splus5/SMB/Allindices.RData")
mri <- 
  All.indices %>% 
  filter(species == 1,
         svaedi == "Heild",
         lengd == 5,
         type == "all") %>% 
  select(year = ar, cb = bio.staerri, cb.cv = cv.bio.staerri, type) %>% 
  mutate(source = "mri")
detach("file:/net/hafkaldi/export/u2/reikn/Splus5/SMB/Allindices.RData")

p +
  geom_pointrange(data = mri, 
                  aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "yellow") +
  geom_line(data = mri, colour = "yellow") +
  labs(x = NULL, y = NULL,
       title = "Biomass indices",
       subtitle = "Comparison of the tidyverse (red) and the Bible (grey)")
```

__Something to worry about__:?
```{r}
tidy_fixed %>% 
  select(year, cb) %>% 
  left_join(mri %>% select(year, cb2 = cb)) %>% 
  mutate(diff = (cb - cb2)/cb2 * 100) %>% 
  summary()
```

So the tidyverse is 0.2% higher than the mri - some may want to dig into that.

#### Grásleppa spring survey index

Here lets try to calculate the index for Grásleppa. We use the same station set and strata as above. Hence only need to do specify in addition the sex:
```{r grasleppa_smb_length, cache=TRUE}
res <- pax::calc_length_indices(Station, Stratas, SPECIES = 48, Sex = 2)
# and a quick plot
res$aggr %>% 
  filter(length == 5) %>% 
  select(year, cb, cb.cv) %>% 
  ggplot(aes(year, cb)) +
  geom_pointrange(aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "red", lwd = 1) +
  geom_line(colour = "red", lwd = 1) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = NULL, y = NULL,
       title = "Grásleppa: Spring survey biomass index",
       subtitle = "All stations")
```

#### "non-standard" species spring survey index

```{r misc_smb_length, cache=TRUE}
spe88 <-  pax::calc_length_indices(Station, Stratas,  88, lwcoeff = c(0.01, 3))
spe562 <- pax::calc_length_indices(Station, Stratas, 562, lwcoeff = c(0.01, 3))
spe150 <- pax::calc_length_indices(Station, Stratas, 150, lwcoeff = c(0.01, 3))
spe71 <-  pax::calc_length_indices(Station, Stratas,  71, lwcoeff = c(0.01, 3))
spe88$aggr %>% 
  mutate(Species = "Rauða sævesla") %>% 
  bind_rows(spe562$aggr %>% mutate(Species = "Ljóskjafta")) %>% 
  bind_rows(spe150$aggr %>% mutate(Species = "Svartgóma")) %>% 
  bind_rows(spe71$aggr %>%  mutate(Species = "Ískóð")) %>% 
  filter(length == 140) %>% 
  select(year, Species, cn, cn.cv) %>% 
  ggplot(aes(year, cn)) +
  geom_pointrange(aes(ymin = cn * (1 - cn.cv),
                      ymax = cn * (1 + cn.cv)),
                  colour = "red", lwd = 1) +
  geom_line(colour = "red", lwd = 1) +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~ Species, scale = "free_y") +
  labs(x = NULL, y = NULL,
       title = "Spring survey abundance index",
       subtitle = "All stations")
```

So we have an invasion of Ljóskjafta and Svartgóma in recent years. That must explain the mackerel invasion :-)

NOTE: There is a bug when it comes to the biomass estimates of the two last species. And one needs to double test for when some species were only counted.

#### And at last be not the least - the fall survey

```{r cod_smh_length, cache=TRUE}
Stations <- 
  pax::smh_strata_setup() %>% 
  select(id = synis.id, year = ar, towlength = toglengd, strata = newstrata) %>% 
  mutate(towlength = pax:::trim_towlength(towlength),
         mult = 1) %>% 
  arrange(id)
Stratas <-
  husky::newstratas_df %>% 
  select(strata, area = rall.area)
res <- pax::calc_length_indices(Stations, Stratas, SPECIES = 1)
res$aggr %>% 
  filter(length == 5) %>% 
  select(year, cb, cb.cv) %>% 
  ggplot(aes(year, cb)) +
  geom_pointrange(aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "red", lwd = 1) +
  geom_line(colour = "red", lwd = 1) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = NULL, y = NULL,
       title = "Cod: Fall survey biomass index",
       subtitle = "All stations")
```


