# Length based indices via mar

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Preamble
___

This case example demonstrates how to calculate standardized survey indices from SMB using the tidyverse approach.

In principle the method is described in [Fjölrit 131](https://www.hafogvatn.is/static/research/files/fjolrit-131pdf). The pseudo-code is something like:

```
1.  get survey stations %>%
2.  get length data %>%
3.  get count data %>% 
4.  scale by counted %>% 
5.  trim towlength %>%                  # narrow extremes
6.  standardize by towlength %>%        # on the number only
7.  calculate_biomass %>%               # up to now one recort per length class
8.  summarise by station %>%            # one record per station
9.  filter stations                     # fixed, e.g. tognumer %in% 1:39
10. summarise by strata %>%             # one record per strata
11. raise to strata area %>%            # 
12. summarise by year                   # one record per year
```

## Setup
___

```{r}
library(tidyverse)
library(mar)
con <- connect_mar()
```


## Additional tables in mar

In order to calculate standardized survey indices within Oracle we need in addition to the tables in mar Oracle tables on:

* Length-weight coefficients
* What strata each tow belongs to
* The area of the strata

For this case example these tables have been made available to all:
```{r}
glimpse(tbl_mar(con, "ops$einarhj.lwcoeff"))
#glimpse(tbl_mar(con, "ops$einarhj.smbstationsstrata"))
glimpse(tbl_mar(con, "ops$einarhj.smb_index_strata"))
glimpse(tbl_mar(con, "ops$einarhj.oldstrataarea"))
```

The smb_index_strata just matches a tow id, consisting of "reitur" and "tognumer" (reitur * 100 + tognnumer) with a particular stratanumber. The oldstrataarea contains the area (in km2) for each strata.

If one is not interested in running the code within Oracle one would need to supply the above tables as an r-dataframe. The side effect, one would need to change the code below such that anytime an Oracle-table is addressed one would need to add the `collect`-function.

A couple of helper functions:
```{r}
#' Calculate overall cv from stratfied summary statistics
#' 
#' @param m Mean value within strata
#' @param s Standard deviation within strata
#' @param area The area (e.g. survey strata area)
#' @param n number of samples within strata
#'
#' @export
#'
calc_cv <- function(m, s, area, n) {
  
  Mean = sum(m * area) / sum(area)
  Sum = sum(m * area)
  tmpsum = sum(m[!is.na(s)] * area[!is.na(s)])
  Calc.sdev = sqrt(sum(s[!is.na(s)]^2 * area[!is.na(s)]^2/  n[!is.na(s)])   / sum(area[!is.na(s)])^2)
  Sdev = Calc.sdev * Sum/tmpsum
  cv = Sdev/Mean
  
  return(cv)
}

# To raise measured fish to the total fish caught
get_count_ratio <- function(lengdir){

  ratio <-
    lesa_numer(lengdir$src) %>%
    dplyr::mutate(r = ifelse(fj_talid==0 | is.na(fj_talid),
                             1,
                             1 + fj_talid / ifelse(fj_maelt == 0 | is.na(fj_maelt), 1, fj_maelt))) %>%
    dplyr::select(synis_id, tegund, r)

  lengdir %>%
    dplyr::left_join(ratio) %>% 
    # for zero stations
    mutate(r = ifelse(is.na(r), 1, r))
}
```


## The code
___

Calculating stratified survey indices is done in four steps:

* A. Data gathering and standardization for each length class
    - Gather the survey, length and counted data for a species of interest.
    - Raise the numbers measured with those counted for each length bin, giving the number of fish caugth in each length bin per station (number of fish caught).
    - Trim extreme tow lengths and then standardize the number of fish caught at each station to a standard towlength (number of fish per standard tow length in each length class).
    - Estimate the weight of fish in each length class from the length-weight coefficient
* B. Summarise abundance and weight for each station 
* C. Calculate mean and standard deviation of abundance and biomass of stations within each strata and raise estimates by the area of the strata
* D. Summarise data by year


Lets select a species to work with and the length above which we want to obtain
the standardized survey indices:
```{r}
Species <- 1                   # Select a species
Length.min <- 5                   # Minimum length for indices calculation
Length.max <- 500                 # Maximum length for indices calculation
```

Some constants used in the standardization process for SMB (you can deviate from them if you want):

```{r}
std.cv        <- 1             # The cv if only one station in a strata
std.towlength <- 4             # Standard tow length for SMB is 4 nautical miles
std.width     <- 17 / 1852     # Standard sweep width in nautical miles

min.towlength <- 2             # Minimum "acceptable" towlength
max.towlength <- 8             # Maximum "acceptable" towlength

Synaflokkur <- 30
Tognumer <- c(1:39, NA)
```


```{r}
# ------------------------------------------------------------------------------
# A. Data gathering

by.length <-  
  # 1. get survey stations -----------------------------------------------------
  lesa_stodvar(con) %>% 
  filter(synaflokkur == Synaflokkur) %>% 
  mutate(index = reitur * 100 + tognumer) %>% 
  select(synis_id, ar, index, reitur, tognumer, veidarfaeri, toglengd) %>% 
  
  # 2. get length data ---------------------------------------------------------
  left_join(lesa_lengdir(con) %>% 
              filter(tegund %in% Species,
                     lengd >= Length.min,
                     lengd < Length.max) %>% 
              group_by(synis_id, tegund, lengd) %>% 
              summarise(fjoldi = sum(fjoldi, na.rm = TRUE)),
            by = "synis_id") %>% 
              ungroup() %>% 
  
  # 0. A temporary fix, for zero stations --------------------------------------
  #     TODO: Find a more permanent solution so scripts works for more than
  #           one species (via group_by( ..., tegund))
  mutate(tegund = if_else(is.na(tegund), Species, tegund),
         lengd  = if_else(is.na(lengd), Length.min, lengd),
         fjoldi = if_else(is.na(fjoldi), 0, fjoldi)) %>% 
  
  # 3. get count data ----------------------------------------------------------
  left_join(lesa_numer(con) %>%
            dplyr::mutate(r = ifelse(fj_talid==0 | is.na(fj_talid),
                                     1,
                                     1 + fj_talid / ifelse(fj_maelt == 0 | is.na(fj_maelt), 1, fj_maelt))) %>%
            dplyr::select(synis_id, tegund, r),
            by = c("synis_id", "tegund")) %>% 
  
  # 4. scale by counted --------------------------------------------------------
  mutate(N = r * fjoldi / 1e3) %>%   # units of thousand
  
  # 5. trim towlength ----------------------------------------------------------
  mutate(toglengd = if_else(toglengd > max.towlength, max.towlength, toglengd),
         toglengd = if_else(toglengd < min.towlength, min.towlength, toglengd)) %>% 
  
  # 6.a standardize by towlength ------------------------------------------------
  mutate(N = N / toglengd * std.towlength) %>%      # standardize to per 4 miles
  # 6.b standardize to area swept
  #     this does not make much sense here because we already have this above
  #     need to pass this function further down in the code path
  mutate(N = N / if_else(veidarfaeri == 78, 1.25 * std.width, std.width)) %>% 
  
  # 7. calculate_biomass from numbers, length and a and b ----------------------
  # 7.a get the length weight coefficients
  left_join(tbl_mar(con, "ops$einarhj.lwcoeff"),
            by = "tegund") %>% 
  # 7.b use Newton's law if lwcoefficient for species not specified
  mutate(a = ifelse(is.na(a), 0.01, a),
         b = ifelse(is.na(b), 3.00, b),
         B  = ifelse(is.na(N), 0, N) * a * lengd^b / 1e3) %>% 
  select(-a, -b) %>% 
  
  # 8. Cumulative calculation
  #     Note: This step is a precursor for calculating things via
  #           year, strata, length
  arrange(tegund, synis_id, lengd) %>% 
  group_by(tegund, synis_id) %>% 
  mutate(cN = cumsum(N),
         cB = sum(B, na.rm = TRUE) - cumsum(B) + B) %>% 
  ungroup()

# ------------------------------------------------------------------------------
# B. Summarise abundance and biomass by station

by.station <- 
  
  by.length %>% 
  
  # 8. summarise by station ----------------------------------------------------
  # NOTE: here is the first step where statistics by length is dropped
  #       some (minor) recoding above would be needed if one were to take things
  #       forward by each length class
  group_by(synis_id, index, reitur, tognumer, veidarfaeri, tegund, ar) %>% 
  summarise(N = sum(N, na.rm = TRUE),
            B = sum(B, na.rm = TRUE)) %>% 
  # Zero stations - THOUGHT THIS STEP SHOULD BE REDUNDANT
  mutate(N = ifelse(is.na(N), 0, N),
         B = ifelse(is.na(B), 0, B))

# ------------------------------------------------------------------------------
# C. Calculate mean and standard deviation of abundance and biomass of stations
#    within each strata and raise estimates by the area of the strata

by.strata <-
  
  by.station %>% 
  
  # 9. filter stations ---------------------------------------------------------  
  filter(tognumer %in% Tognumer) %>%
  
  # 10. summarise by strata ----------------------------------------------------
  # 10.a  Get the strata for each station
  left_join(tbl_mar(con, "ops$einarhj.smb_index_strata") %>% 
              select(index, strata = stdoldstrata)) %>% 
  # 10.b group by year and strata and calculate number of stations, mean and sd
  group_by(tegund, ar, strata) %>% 
  summarise(sN  = n(),   # number of stations within strata
            n_m  = mean(N, na.rm = TRUE),
            n_d  = ifelse(n() == 1, mean(N, na.rm = TRUE) * std.cv, sd(N)),
            b_m  = mean(B, na.rm = TRUE),
            b_d  = ifelse(n() == 1, mean(B, na.rm = TRUE) * std.cv, sd(B))) %>% 
  
  # 11. raise to strata area ---------------------------------------------------
  # 11.a get area of the strata
  left_join(tbl_mar(con, "ops$einarhj.oldstrataarea") %>% 
              select(strata = oldstrata, area = rall.area) %>% 
              #  area is above is in km2, here convert nm2
              mutate(area = area / 1.852^2)) %>% 
  # 11.b do the strata raising
  mutate(n     = n_m  * area / std.towlength,
         b     = b_m  * area / std.towlength) 

# ------------------------------------------------------------------------------
# D. Summarise data by year

by.year <- 
  
  by.strata %>% 
  
  
  # ----------------------------------------------------------------------------
  # Up to now we only have been operating within Oracle. I.e. sql-scripts via R.
  # Have to collect here because of calc_cv function in the year aggregate step
  # TODO: Fix that, do internally in Oracle
  collect(n = Inf) %>% 

  
  # ----------------------------------------------------------------------------
  # some data fall outside strata, drop them - needs some double checking of code
  drop_na() %>% 
  
  # 11. summarise by year ------------------------------------------------------
  group_by(tegund, ar) %>%
  summarise(n = sum(n, na.rm = TRUE),
            # A la Höski
            n.cv = calc_cv(n_m, n_d, area, sN),
            b = sum(b, na.rm = TRUE),
            # A la Höski
            b.cv = calc_cv(b_m, b_d, area, sN)) %>% 
  ungroup()
```



## The results
___

```{r}
glimpse(by.year)
```

So we have a dataframe where each row is year and then the abundance and biomass indices and the cv. Lets plot the biomass indices:
```{r}
by.year %>% 
  ggplot(aes(ar + 3/12, b)) +
  geom_point() +
  geom_linerange(aes(ymin = b * (1 - b.cv),
                      ymax = b * (1 + b.cv))) +
  expand_limits(y = 0) +
  labs(x = NULL, y = NULL,
       title = "Spring survey biomass indices") +
  facet_wrap(~ tegund, scale = "free_y")
```


## Meeting with the Pope

Below a comparison with results based on Pope Husky's code, just as a double check.

```{r, echo = FALSE}
attach("/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Allaggroldsmbindex.rdata")
vatican <-
 Allaggroldsmbindex %>%
 filter(species == Species,
        svaedi == "Heild",
        diurnal == 0,
        fixed == 0) %>%
 select(ar, lengd, b = bio.staerri, b.cv = cv.bio.staerri) %>%
 mutate(source = "Vatican") %>%
 as_tibble()
vatican <-
  vatican %>% 
  filter(lengd == max(min(lengd), Length.min))
detach("file:/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Allaggroldsmbindex.rdata")

by.year %>% 
  mutate(source = "tidy",
         ar = ar + 3/12) %>% 
  bind_rows(vatican) %>% 
  ggplot(aes(ar, b, colour = source)) +
  geom_point() +
  geom_pointrange(aes(ymin = b * (1 - b.cv),
                      ymax = b * (1 + b.cv))) +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.position = c(0.25, 0.8)) +
  expand_limits(y = 0) +
  labs(x = NULL, y = NULL,
       title = "Biomass indices",
       subtitle = "Comparison of the orthodoxy (Vatican) and tidy")
```

Besides that tidy and the Vatican are not operating in the same time zone (one has argued that tidy operates under the Greek-orthodox time zone, but it seems inverse) the stuff on the y-axis looks reasonably similar.

## Older stuff

__The flow__:
```
Stations %>%
trim towlength %>%                 # narrow extremes
gather length data %>%
scale by counted %>% 
standardize by towlength %>%       # on the number only
calculate_biomass %>%              # up to now on each length class
summarise_by_station %>%           # now one record per station
                                   # Note statistic up to now is equivalent
                                   #   to afli per tog
                                   # HERE: filter station to use
                                   #       i.e. fixed etc
summarise_by_strata %>%            # one record per strata
summarise                          # one record per year
```

__Loading needed libraries__:

```{r length_indices_mar}
library(tidyverse)
library(mar)
```

__Some functions__:
```{r}
trim_towlength <- function(d, min.towlength = 2, max.towlength = 8) {
  
  d %>% 
    mutate(toglengd = if_else(toglengd > max.towlength, max.towlength, toglengd),
           toglengd = if_else(toglengd < min.towlength, min.towlength, toglengd))
}


gather_length_data <- function(Stations, Species, Length = 0) {
  
  # Here could create an switch, if Stations is connection do below
  #   else read from fjolst
  Lengths <- 
    lesa_lengdir(Stations$src) %>% 
    filter(tegund %in% Species,
           lengd >= Length) %>% 
    group_by(synis_id, tegund, lengd) %>% 
    summarise(fjoldi = sum(fjoldi, na.rm = TRUE)) %>% 
    ungroup()
  
  Counts <-
    lesa_numer(Stations$src) %>% 
    filter(tegund %in% Species) %>% 
    mutate(r =  ifelse(fj_maelt != 0, 1 + fj_talid/fj_maelt, 1)) %>% 
    select(synis_id, tegund, fj_maelt, fj_talid, r)
  
  Stations %>% 
    left_join(Lengths, by = "synis_id") %>% 
    left_join(Counts, by = c("synis_id", "tegund"))
  
}


scale_by_counted <- function(d) {
  d %>% 
    mutate(fjoldi = r * fjoldi / 1e3) %>%   # units of thousand
    select(-r)                              # no use anymore
}

standardise_towlength <- function(d, std.towlength = 4) {
  
  d %>% 
    mutate(N = fjoldi * std.towlength/toglengd) %>%   # standardize to per 4 miles
    select(-toglengd)

}

calculate_biomass <- function(d) {
  
  lwcoeff <- tbl_mar(con, "ops$einarhj.lwcoeff")
  
  d %>% 
    left_join(lwcoeff, by = "tegund") %>% 
    # use Newton's law if lwcoefficient for species not specified
    mutate(a = ifelse(is.na(a), 0.01, a),
           b = ifelse(is.na(b), 3.00, b)) %>% 
    # Below statistic is by length class
    # If include lwcoeff then do a priori a left_join
    mutate(B  = ifelse(is.na(N), 0, N) * a * lengd^b/1e3) %>% 
    select(-a, -b)
  
}

calc_by_strata <- function(d) {
  
  # Stuff should not really be inside this function
  Stratas <- 
  tbl_mar(con, "ops$einarhj.stratas") %>% 
  select(strata, area = rall.area)
    
  d %>% 
    group_by(ar, strata) %>% 
  dplyr::summarise(sN  = n(),   # Number of stations within strata
                   n_m  = mean(N, na.rm = TRUE),
                   n_d  = ifelse(n() == 1, mean(N, na.rm = TRUE) * std.cv, sd(N)),
                   b_m  = mean(B, na.rm = TRUE),
                   b_d  = ifelse(n() == 1, mean(B, na.rm = TRUE) * std.cv, sd(B))) %>% 
  #b_d  = ifelse(N == 1, b_m  * std.cv, stats::sd(b)),
  ungroup() %>%
  left_join(Stratas, by = "strata") %>%
  # within oracle, have to mutate in stages
  mutate(area  = area/1.852^2 / std.area) %>% 
  mutate(n     = n_m  * area,
         b     = b_m  * area)
  
}

aggregate_by_year <- function(d) {
  
  # below equvialent to aggr-calculation
  d %>% 
    group_by(ar) %>%
    # A la Höski:
    summarise(n = sum(n, na.rm = TRUE),
              n.cv = pax:::calc_cv(n_m, n_d, area, sN),
              b = sum(b, na.rm = TRUE),
              b.cv = pax:::calc_cv(b_m, b_d, area, sN)) %>% 
    ungroup()
  
}
```

__Run the flow__:
```{r}
LENGTH <- 5 # Index for fish length equal to or larger than
con <- connect_mar()
```

**Get the oldstrata for each smb and store it in Oracle**:
```{r, eval = FALSE}
# NOT RUN
attach("/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Stations.rdata")
smb <- 
  STODVAR.all %>%
  select(synis_id = synis.id, ar, reitur, tognumer, toglengd, strata = oldstrata, lon, lat)
dbWriteTable(con, name = "smbstations", value = smb, overwrite = TRUE)
detach("file:/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Stations.rdata")
```


```{r}
Stations <-  
  tbl_mar(con, "ops$einarhj.smbstations")
std.cv <- 1
std.area <- 4 * 17/1852
Stations %>%
  trim_towlength() %>% 
  #left_join(Stratas) %>% 
  gather_length_data(Species = 1, Length = LENGTH) %>%
  scale_by_counted() %>% 
  standardise_towlength() %>% 
  calculate_biomass() ->
  # NOTE: upto here we have data by tow and length
  #       hence break the chain here if one were intersted
  #       in catch per station
  std_catch_per_station_per_length

# SIDEKICK: Now we could do:
std_catch_per_station_per_length %>% 
  group_by(ar, lon, lat) %>% 
  summarise(N = sum(N, na.rm = TRUE),
            B = sum(B, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(ar %in% c(1985, 1995, 2005, 2015)) %>% 
  collect(n = Inf) %>% 
  ggplot(aes(lon, lat, size = B)) +
  geom_point(col = "red", alpha = 0.4) +
  scale_size_area(max_size = 20) +
  facet_wrap(~ ar) +
  coord_quickmap() +
  labs(x = NULL, y = NULL, size = "Catch [t]",
       title = "Catch of cod",
       subtitle = "'Standardized' towlength")

# Here proceed with the main objectives:
std_catch_per_station_per_length %>% 
  # NOTE: next step is the first filtering of stations
  filter(tognumer %in% 1:39 | is.na(tognumer)) %>%
  group_by(synis_id, ar, strata, tegund) %>% 
  summarise(N = sum(N, na.rm = TRUE),
            B = sum(B, na.rm = TRUE)) %>% 
  # Zero stations
  mutate(N = ifelse(is.na(N), 0, N),
         B = ifelse(is.na(B), 0, B)) %>% 
  calc_by_strata() %>% 
  # Have to collect here because of calc_cv function in the aggregate step
  collect(n = Inf) %>% 
  aggregate_by_year() ->
  d
```

__A double test__:
```{r}
attach("/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Allaggroldsmbindex.rdata")
vatican <- 
  Allaggroldsmbindex %>% 
  filter(species == 1,
         svaedi == "Heild",
         lengd == LENGTH,
         diurnal == 0,
         fixed == 0) %>% 
  select(year = ar, cb = bio.staerri, cb.cv = cv.bio.staerri) %>% 
  mutate(source = "vatican") %>% 
  as_tibble()
detach("file:/net/hafkaldi/export/u2/reikn/R/SurveyWork/SMB/Allaggroldsmbindex.rdata")

d %>% 
  select(year = ar, cb = b, cb.cv = b.cv) %>% 
  mutate(source = "tidy",
         year = year + 0.2) %>% 
  bind_rows(vatican) %>% 
  ggplot(aes(year, cb)) +
  geom_pointrange(aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv),
                      colour = source),
                  lwd = 1) +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.position = c(0.1, 0.8)) +
  labs(x = NULL, y = NULL,
       title = "Biomass indices",
       subtitle = "Comparison of the official (vatican) and tidy")
```
