# On stations

## SMB

In 2017 the tow-index for tows having "tognumer" less than 40 have been fixed to strata. The code used in 2018 to generate the "official" index can be found under /u2/reikn/R/SurveyWork/SMB/stodvar.r

What follows is a tidy mimicry of that code:

```{r, message = FALSE, warning = FALSE}
library(mar)
con <- connect_mar()

index.strata <- 
  tbl(con, "SMB_INDEX_STRATA_AREA")

st <-
  lesa_stodvar(con) %>%
  filter(synaflokkur == 30) %>%
  mutate(tognumer = ifelse(synis_id == 474396, 1, tognumer),
         index = reitur * 100 + tognumer,
         # not needed
         # filter(!index %in% c(56835,56836))
         lon = ifelse(is.na(hift_v_lengd),
                      kastad_v_lengd,
                      (kastad_v_lengd + hift_v_lengd) / 2),
         lat = ifelse(is.na(hift_n_breidd),
                      kastad_n_breidd,
                      (kastad_n_breidd + hift_n_breidd) / 2)) %>% 
  left_join(index.strata, by = "index")
# checking - all townumbers less than 40 should have a defined strata
st %>% 
  mutate(less40 = ifelse(tognumer < 40, 1, 0),
         def.old = ifelse(!is.na(stdoldstrata), 1, 0),
         def.new = ifelse(!is.na(stdnewstrata), 1, 0)) %>% 
  group_by(less40) %>% 
  summarise(n = n(),
            def.old = sum(def.old, na.rm = TRUE),
            def.new = sum(def.new, na.rm = TRUE)) %>% 
  collect(n = Inf)
# So all tows (19142) less than 40 have assigned a strata
#    123 tows are above 40 and they do not (naturally have a strata)

# function in package husky changed so it returns a vector
inside_strata <- 
  function (stodvar, stratalist = ralllist, stratas = STRATAS) {
  stodvar$newstrata <- rep(0, nrow(stodvar))
  for (i in stratalist) {
    j <- geo::geoinside(stodvar, reg = stratas[[i]], robust = T, 
                        option = 0)
    if (length(j) > 0) 
      stodvar[j, "newstrata"] <- i
  }
  i <- stodvar$newstrata == 0
  if (any(i)) {
    i1 <- 1:length(i)
    i1 <- i1[i]
    cat("Strata fannst ekki fyrir stodvar", paste(i1, collapse = ","))
  }
  return(stodvar$newstrata)
}

# ------------------------------------------------------------------------------
# For stations > 40
d <-
  st %>% 
  collect(n = Inf)

# load STRATAS
# Rattach("/u2/reikn/R/SurveyWork/OldStratas", pos = 2)
d <- 
  d %>%
  mutate(oldstrata = inside_strata(.,
                                   stratalist = husky::ralllist,
                                   stratas = husky::STRATAS),
         newstrata = inside_strata(.,
                                   stratalist = husky::smblist$nr,
                                   stratas = husky::NEWSTRATAS))
# Smá lagfæringar
d <- 
  d %>% 
  mutate(oldstrata = ifelse(oldstrata %in% 23, 88, oldstrata),
         newstrata = ifelse(index %in% 71741 & ar %in% 2009, 37, newstrata),
         newstrata = ifelse(index %in% 61141 & ar %in% 2010, 42, newstrata))

d <-
  d %>% 
  mutate(oldstrata = ifelse(tognumer < 40, stdoldstrata, oldstrata),
         newstrata = ifelse(tognumer < 40, stdnewstrata, newstrata)) %>% 
  select(-c(stdoldstrata, stdnewstrata))

# ------------------------------------------------------------------------------
# checking
library(tidyverse)
load("/u2/reikn/R/SurveyWork/SMB/Stations.rdata")
STODVAR.all %>% 
  filter(tognumer < 40) %>% 
  select(index, area) %>% 
  distinct() %>% 
  group_by(index) %>% 
  count() %>% 
  arrange(-n)

d2 <- STODVAR.all

x <- 
  d %>% 
  select(synis_id, index, oldstrata, newstrata) %>% 
  arrange(synis_id)
x2 <-
  STODVAR.all %>%
  select(synis_id = synis.id, index, oldstrata, newstrata) %>% 
  as_tibble() %>% 
  arrange(synis_id)
identical(x, x2)
```

Now check for the Bormicon area
```{r}
x <- 
  d %>%
  filter(tognumer < 40) %>% 
  select(synis_id, area) %>% 
  arrange(synis_id)
x2 <-
  STODVAR.all %>% 
  filter(tognumer < 40) %>% 
  select(synis_id = synis.id, area) %>% 
  as_tibble() %>% 
  arrange(synis_id)
identical(x, x2)

STODVAR.all %>% 
  filter(tognumer < 40) %>% 
  select(index, area) %>% 
  distinct() %>% 
  group_by(index) %>% 
  count() %>% 
  arrange(-n)
```

So we have in the official calculation index that falls in different bormicon areas in different years. Not something I expected.

Lets see however if we can mimic this for the tidyverse:
```{r}
library(geo)
x <- 
  d %>% 
  filter(tognumer < 40) %>% 
  select(synis_id, index, lon, lat, area.fixed = area)
x <- inside.reg.bc(x)
x %>% filter(area.fixed != area)
x <- x %>% select(synis_id, area) %>% arrange(synis_id)
identical(x, x2)
```

Lets try for all stations, i.e. also those over 40
```{r}
x <- 
  d %>% 
  select(synis_id, index, lon, lat, area.fixed = area)
x <- inside.reg.bc(x)
x <- x %>% select(synis_id, area) %>% arrange(synis_id)
x2 <-
  STODVAR.all %>% 
  select(synis_id = synis.id, area) %>% 
  as_tibble() %>% 
  arrange(synis_id)
identical(x, x2)
```

## Start from scratch

```{r}
library(mar)
con <- connect_mar()

st <-
  lesa_stodvar(con) %>%
  filter(synaflokkur == 30) %>%
  mutate(tognumer = ifelse(synis_id == 474396, 1, tognumer)) %>%
  mutate(index = reitur * 100 + tognumer) %>%
  # not needed
  # filter(!index %in% c(56835,56836))
  mutate(lon = ifelse(is.na(hift_v_lengd),
                      kastad_v_lengd,
                      (kastad_v_lengd + hift_v_lengd) / 2),
         lat = ifelse(is.na(hift_n_breidd),
                      kastad_n_breidd,
                      (kastad_n_breidd + hift_n_breidd) / 2)) %>% 
  left_join(tbl(con, "SMB_INDEX_STRATA")) %>% 
  collect(n = Inf) %>% 
  mutate(oldstrata = inside_strata(.,
                                   stratalist = husky::ralllist,
                                   stratas = husky::STRATAS),
         newstrata = inside_strata(.,
                                   stratalist = husky::smblist$nr,
                                   stratas = husky::NEWSTRATAS)) %>% 
  # Some fixes -----------------------------------------------------------------
  mutate(oldstrata = ifelse(oldstrata %in% 23, 88, oldstrata),
         newstrata = ifelse(index %in% 71741 & ar %in% 2009, 37, newstrata),
         newstrata = ifelse(index %in% 61141 & ar %in% 2010, 42, newstrata)) %>% 
  # Use fixed strata for "tognumer" less than 40 -------------------------------
  mutate(oldstrata = ifelse(tognumer < 40, stdoldstrata, oldstrata),
         newstrata = ifelse(tognumer < 40, stdnewstrata, newstrata)) %>% 
  select(-c(stdoldstrata, stdnewstrata)) %>% 
  # Get the bormicon area for the stations -------------------------------------
  inside.reg.bc(.)
```

Double checking:
```{r}
x <-
  st %>% 
  select(synis_id, oldstrata, newstrata, area) %>% 
  arrange(synis_id)
x2 <-
  STODVAR.all %>% 
  select(synis_id = synis.id, oldstrata, newstrata, area) %>% 
  as_tibble() %>% 
  arrange(synis_id)
identical(x, x2)
```


## Trial with geo_inside and shapefiles

```{r}
library(mar)
con <- connect_mar()
library(gisland)
st <-
  lesa_stodvar(con) %>%
  filter(synaflokkur == 30) %>%
  mutate(tognumer = ifelse(synis_id == 474396, 1, tognumer)) %>%
  mutate(index = reitur * 100 + tognumer) %>%
  # not needed
  # filter(!index %in% c(56835,56836))
  mutate(lon = ifelse(is.na(hift_v_lengd),
                      kastad_v_lengd,
                      (kastad_v_lengd + hift_v_lengd) / 2),
         lat = ifelse(is.na(hift_n_breidd),
                      kastad_n_breidd,
                      (kastad_n_breidd + hift_n_breidd) / 2)) %>% 
  left_join(tbl(con, "SMB_INDEX_STRATA")) %>% 
  collect(n = Inf) %>% 
  mutate(oldstrata = geo_inside(lon, lat, gisland::stratas.sp, variable = "ID"),
         newstrata = geo_inside(lon, lat, gisland::stratasnew.sp, variable = "ID")) %>% 
  # Some fixes -----------------------------------------------------------------
  mutate(oldstrata = ifelse(oldstrata %in% 23, 88, oldstrata),
         newstrata = ifelse(index %in% 71741 & ar %in% 2009, 37, newstrata),
         newstrata = ifelse(index %in% 61141 & ar %in% 2010, 42, newstrata)) %>% 
  # Use fixed strata for "tognumer" less than 40 -------------------------------
  mutate(oldstrata = ifelse(tognumer < 40, stdoldstrata, oldstrata),
         newstrata = ifelse(tognumer < 40, stdnewstrata, newstrata)) %>% 
  select(-c(stdoldstrata, stdnewstrata)) %>% 
  # Get the bormicon area for the stations -------------------------------------
  mutate(area = geo_inside(lon, lat, gisland::bormicon, variable = "ID"))
```

Double checking:
```{r}
x <-
  st %>% 
  select(synis_id, oldstrata, newstrata, area) %>% 
  arrange(synis_id)
x2 <-
  STODVAR.all %>% 
  select(synis_id = synis.id, lon2 = lon, lat2 = lat, oldstrata, newstrata, area) %>% 
  as_tibble() %>% 
  arrange(synis_id)
identical(x, x2)
identical(x$synis_id, x2$synis_id)
identical(as.integer(x$area), as.integer(x2$area))
identical(as.integer(x$oldstrata), as.integer(x2$oldstrata))
identical(as.integer(x$newstrata), as.integer(x2$newstrata))
```

Seem to be boarderline cases:

```{r}
st %>% 
  select(synis_id, index, o1 = oldstrata) %>% 
  full_join(x2 %>% select(synis_id, o2 = oldstrata)) %>% 
  filter(o1 != o2) %>% 
  knitr::kable()
tmp <- 
  st %>% 
  select(synis_id, lon1 = lon, lat1 = lat, n1 = newstrata) %>% 
  full_join(x2 %>% select(synis_id, lon2, lat2, n2 = newstrata)) %>% 
  filter(n1 != n2)
tmp %>% 
  knitr::kable()
near(tmp$lon1, tmp$lon2) %>% sum()
table(tmp$n1, tmp$n2)
```

```{r}
tmp2 <- tmp %>% select(long = lon1, lat = lat1)
library(leaflet)
leaflet(stratasnew.sp) %>% 
  addTiles() %>% 
  addPolygons() %>% 
  addCircleMarkers(data = tmp2, col = "red")
```

